Is the Future of Authenticity In Our Heads?\\\medskip \large Moving Passthoughts From the Lab to the World

### 2017-04-14 authors.tex
revisions

\begin{abstract}
\textit{Passthoughts authentication}, in which a user thinks a secret thought to log in to services or devices,
allows for two factors of authentication (knowledge and inherence) in a single step.
Since it was first proposed in 2005, passthoughts has seen a number of successful empirical studies.
In this paper, we renew and justify the importance and promise of passthoughts authentication.
At the same time, we review the main challenges passthoughts must overcome in order to move from the lab to the real world.
We propose two studies that could make significant progress on this front,
both of which seek different angles around the fundamental questions we pose.
We conclude with future oppportunities, and challenges, for passthoughts' broader deployment as a tool for authentication.
In so doing, we raise novel possibilities for authentication more broadly,
such as ``organic passwords'' that change naturally over time,
or systems that reject users who are not acting quite like ``themselves.''
\end{abstract}

>>>>

\begin{abstract}
\textit{Passthoughts authentication}, in which a user thinks a secret thought to log in to services or devices,
provides two factors of authentication (knowledge and inherence) in a single step.
Since its proposal in 2005, passthoughts has seen a number of successful empirical studies.
In this paper, we renew and justify the promise of passthoughts authentication,
reviewing the main challenges that passthoughts must overcome in order to move from the lab to the real world.
We propose two studies,
both of which seek different angles at the fundamental questions we pose.
We conclude with future oppportunities, and challenges, for passthoughts' broader deployment as a tool for authentication.
In so doing, we raise novel possibilities for authentication more broadly,
such as ``organic passwords'' that change naturally over time,
or systems that reject users who are not acting quite like ``themselves.''
\end{abstract}

### 2017-04-14 paper.tex
revisions

% Created 2017-04-13 Thu 17:24
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\usepackage{lmodern}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.0.50 (Org mode 9.0.5)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:orgdc66532}

@@

While recent work is heartening, passthoughts remains confined, for now, to the lab.
This paper reviews the immediate questions that must be addressed if passthoughts is to move into everyday life.
First, we must understand the real usability properties of passthoughts.
While much work has claimed passthoughts to be more usable than traditional multi-factor authentication protocols,
as of yet, no work systematically evaluates this claim.
The usability of passthoughts authentication will depend not only on its performance in ecologically valid contexts,
but also on people's attitudes about brainwaves, EEG, and biosensing generally (Section 3).
Second, we must test passthoughts in a variety of conditions: ambulatory settings, under different levels of stress, drowsiness, caffeine or alcohol, etc.
This statistical analysis will help us understand how passthoughts change over time.
It will also help us understand the space of possible passthoughts (Section 4).

@@

The main goal of this paper is to motivate and contextualize studies that can begin to address these challenges (Section 5).
The first study focuses on the use of passthoughts in a realistic, though lab-constrained setting, 
allowing a first look at how passthoughts might work in a real-time application context.
The second proposed study focuses on people's longitudinal relationships with EEG in their everyday life,
focusing not on behavioral differences experimental conditions, but on evolving attitudes (and signals) over longer timescales,
allowing assessment of both dynamic user attitudes, and the possible impact of shifting neural signals.

Passthought authentication stands a chance at becoming the first brain-computer interface to reach wider adoption. 
As such, passthoughts promises not only more usable multi-factor authentication,
but also a source of data for future development of EEG-based brain-computer interfaces
e.g. for people with motor disabilities \cite{Mattia2013}.
Passthoughts is not without risks (Section 6),
but, if future work can address the two challenges posed here, several interesting paths unfurl.
closed-loop authentication, continuous and/or tacit authentication, 
and possibly novel contributions to authentication more broadly (Section 7).

\section{Background}
\label{sec:org86b04e6}

Authentication is the act of proving that a user is who they claim to be.
In computer security, authenticators are classified into three types: knowledge factors (e.g., passwords
and PINs), possession factors (e.g., physical tokens, ATM cards), and inherence
factors (e.g., fingerprints and other biometrics).

@@

\subsection{One-step, multi-factor authentication}
\label{sec:orgb8e9707}

@@

\subsection{Passthought authentication}
\label{sec:orgca3bf6a}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:org72b8e9d}

@@

\section{User attitudes and perceptions}
\label{sec:orgec86f2d}

@@

\subsection{Contending with mind-reading machines}
\label{sec:org764904d}

@@

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:orgea43368}

@@

\section{Diversity and security of passthoughts}
\label{sec:orgd9c53c7}

@@

\section{Two studies on passthoughts}
\label{sec:org354b38d}

@@

\subsection{A real-time passthought authenticator}
\label{sec:orga45ff6f}

@@

\subsubsection{Study protocol}
\label{sec:org7ffd699}

@@

\subsubsection{The effect of feedback}
\label{sec:org5931401}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:org54e98fb}

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:org64b1300}

@@

\subsubsection{Study protocol}
\label{sec:org07b7e0e}
A small group of subjects could wear a working, recording EEG device, whether or not it provides feedback, in a variety of settings for some number of days,
having subjects journal their experiences and asking them specifically what they feel someone might be able to know about them from the EEG signals they record.
At the same time, we could use this study as an opportunity to collect a much larger, and more diverse corpus.
To aid in the collection of signals that are specific to our problem of passthought authentication,
subjects in this study might be prompted to perform a variety of tasks at a few pre-chosen points throughout the day.
With the data collected during this study, we could easily simulate passthought accuracy on a much more realistic (and representative) sample of readings.

@@

\subsubsection{A more diverse corpus}
\label{sec:orgd9d792a}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:orgbee77bb}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how rare or unlikely a given passthought is, and help shed light on why and how given passthoughts are expressed uniquely between individuals.

@@

\subsubsection{Usability and attitudes}
\label{sec:orge4bf59b}

@@

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:org8b26c17}

@@

\subsection{Privacy}
\label{sec:orgc4b91c1}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:org3990614}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

@@

\subsection{Tradeoffs between security and privacy}
\label{sec:org41bf0f2}

@@

\section{Further Future Directions}
\label{sec:org663b329}

@@

\subsection{Closed-loop (real-time) passthoughts}
\label{sec:org36e36f1}
Future work on passthoughts should look at closed-loop, or online authentication systems,
in part to investigate the impact of human learning effects on passthought performance.
What effect does the feedback (of a successful or unsuccessful authentication attempt) have on the way that people perform their passthoughts?
Specific studies could, for example, provide false feedback in which passthought authentication appears to always either succeed or fail. In the always-fail condition, we might expect subjects to alter the way they perform the passthought across multiple attempts; data of how such a change occurs could enable us to preempt changes observable in the wild.

\subsection{Continuous authentication}
\label{sec:org0209ad5}

@@

\subsection{Organic passwords}
\label{sec:orgd47e984}

If EEG signals are nonstationary (changing over time), passthoughts will require continuous re-calibration to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as voice or keystroke dynamics \cite{Monrose1997}, which may change more slowly for individuals, if they change at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Neuroscience of authentication}
\label{sec:org250810d}

@@

Such a scenario raises serious ethical, legal, and even philosophical questions. 
How does such a system conform to accepted definitions of a ``person''?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as I am aware, novel, and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this still-novel method for authentication.
\subsection{Passthoughts by any other sensor?}
\label{sec:org65e1ff9}

@@

\subsection{Health, neuroscience and BCIs}
\label{sec:org3216f65}

@@

\section{Conclusion}
\label{sec:org7610c05}

>>>>

% Created 2017-04-13 Thu 18:06
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\usepackage{lmodern}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.0.50 (Org mode 9.0.5)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:org0a430e5}

@@

@@

While recent work is heartening, passthoughts remains confined, for now, to the lab.
This paper reviews the immediate questions that must be addressed if passthoughts is to move into everyday life.
First, we must understand the real usability properties of passthoughts.
While much work has claimed passthoughts to be more usable than traditional multi-factor authentication protocols,
as of yet, no work systematically evaluates this claim.
The usability of passthoughts authentication will depend not only on its performance in ecologically valid contexts,
but also on people's attitudes about brainwaves, EEG, and biosensing generally (Section 3).
Second, we must test passthoughts in a variety of conditions: ambulatory settings, under different levels of stress, drowsiness, caffeine or alcohol, etc.
This statistical analysis will help us understand 
the space of possible passthoughts,
and how passthoughts change over time (Section 4).

This paper motivates and contextualizes two studies that could begin to address these challenges (Section 5).
While these studies attack two sides of passthoughts' chances as a real-world technology
this study serves not just to propose solutions, but to raise questions about what authentication could and should be able to do.
While passthoughts is not without risks (Section 6),
it raises the possiblity for several interesting paths,
including exploration of systems that could enable novel types of authentication (Section 7).

\section{Background}
\label{sec:orgc5b76aa}

Authentication seeks to prove that a user is who they claim to be.
In computer security, authenticators are classified into three types: knowledge factors (e.g., passwords
and PINs), possession factors (e.g., physical tokens, ATM cards), and inherence
factors (e.g., fingerprints and other biometrics).

@@

\subsection{One-step, multi-factor authentication}
\label{sec:orgce77dd1}

@@

\subsection{Passthought authentication}
\label{sec:orgddf4083}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:org7c65df9}

@@

\section{User attitudes and perceptions}
\label{sec:org1e82386}

@@

\subsection{Contending with mind-reading machines}
\label{sec:org6b82b47}

@@

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:orgb912760}

@@

\section{Diversity and security of passthoughts}
\label{sec:orga4f01a7}

@@

\section{Two studies on passthoughts}
\label{sec:orgda992dc}

@@

\subsection{A real-time passthought authenticator}
\label{sec:org647a2bd}

@@

\subsubsection{Study protocol}
\label{sec:org5f58e5b}

@@

\subsubsection{The effect of feedback}
\label{sec:org68a3874}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:org236c4c5}

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:orgdb1427a}

@@

\subsubsection{Study protocol}
\label{sec:org0ed08a7}
A small group of subjects could wear a working, recording EEG device, whether or not it provides feedback, in a variety of settings for some number of days,
having subjects journal their experiences and asking them specifically what they feel someone might be able to know about them from the EEG signals they record.
At the same time, we could use this study as an opportunity to collect a much larger, and more diverse corpus.
To aid in the collection of signals that are specific to our problem of passthought authentication,
subjects in this study might be prompted to perform a variety of tasks at a few pre-chosen points throughout the day.
With the data collected during this study, we could easily simulate passthought accuracy on a much more realistic (and representative) sample of readings.

@@

\subsubsection{A more diverse corpus}
\label{sec:org5eb9cdc}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:org40f3dce}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how rare or unlikely a given passthought is, and help shed light on why and how given passthoughts are expressed uniquely between individuals.

@@

\subsubsection{Usability and attitudes}
\label{sec:orgd44d815}

@@

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:org7e5111c}

@@

\subsection{Privacy}
\label{sec:org8a9010b}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:org22dc2d3}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

@@

\subsection{Tradeoffs between security and privacy}
\label{sec:org9627bbb}

@@

\section{Further Future Directions}
\label{sec:orgd918e94}

@@

\subsection{Closed-loop (real-time) passthoughts}
\label{sec:org80cf40b}
Future work on passthoughts should look at closed-loop, or online authentication systems,
in part to investigate the impact of human learning effects on passthought performance.
What effect does the feedback (of a successful or unsuccessful authentication attempt) have on the way that people perform their passthoughts?
Specific studies could, for example, provide false feedback in which passthought authentication appears to always either succeed or fail. In the always-fail condition, we might expect subjects to alter the way they perform the passthought across multiple attempts; data of how such a change occurs could enable us to preempt changes observable in the wild.

\subsection{Continuous authentication}
\label{sec:org7d58379}

@@

\subsection{Organic passwords}
\label{sec:orgc942193}

If EEG signals are nonstationary (changing over time), passthoughts will require continuous re-calibration to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as voice or keystroke dynamics \cite{Monrose1997}, which may change more slowly for individuals, if they change at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Neuroscience of authentication}
\label{sec:orgda2b5f8}

@@

Such a scenario raises serious ethical, legal, and even philosophical questions. 
How does such a system conform to accepted definitions of a ``person''?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as I am aware, novel, and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this still-novel method for authentication.
\subsection{Passthoughts by any other sensor?}
\label{sec:org495f530}

@@

\subsection{Health, neuroscience and BCIs}
\label{sec:org70dd836}

@@

\section{Conclusion}
\label{sec:orga01fe32}
